---
title: "Data Extraction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data Extraction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

For the algorithm to work, you need to extract from your cut of MarketScan (Commercial Claims or Medicaid) specific claims records (marked by the codes stored in the codelists) for services between specific dates on women within specific age boundaries. This is the first step highlighted in the [pregfindr vignette](pregfindr.html).

The result of this extraction should follow the following format:
```{r eval=FALSE, echo = TRUE}
?pregfindr::extraction_example
```

Since different companies use different technologies to store the latest cut of MarketScan, we are sharing an example script that you can customise for your data storage system.
This script uses helper functions defined in the package and specific codelists. To learn more about the codelists and how to modify them, refer to the [codelist vignette](codelists.html).

```{r eval=FALSE, echo=TRUE}
# The script is designed to extract, process, and analyse healthcare data from the
# MarketScan database. It identifies eligible women within a specific age range and study
# period, retrieves their medical claims data, matches these with
# pregnancy-related diagnosis and procedure codes, and combines all the data
# into a single dataset. The script also includes logging and data saving
# functionalities.

# Loading packages
library(pregfindr)
library(logr)
library(DBI)
library(dplyr)
library(dbplyr)
library(lubridate)

# Starting log -------------
log_file <- log_open(show_notes = FALSE)

# Setting variables ------------
sep("Variables")

# Pointer to latest cut of MarketScan available and key parameters for analysis
cut <- "2023q2"

# Set Study Time Period - time frame of analysis
study_start_date <- "01Jan2021" # study start date
service_start_date <- "2021-01-01"

# when pregnancy ends outside of range, it will have unknown outcome
study_end_date <- "31Dec2021" # study end date
service_end_date <- "2021-12-31"
start_yr <- 2021 # Study start year
last_yr <- 2021 # Study end year

# pointer to key parameters for analysis - when drugs are approved for adult patients, then we might want to subset
age_start <- 30 # ages of interest
age_end <- 60 # ages of interest
mem_days <- 1 # minimum follow-up in a year - since want to find all who ever use meds
# typically set to one day

# Number of days before study start to assess pregnancy codes - full term pregnancies are 270 days - allows for longer pregnancies
presvcstart <- 300

# Number of days for the pre-conception period - to allow for drugs to be flushed out
pre_conception_days <- 90

# Number of DAYS of required enrolment before and after index  - to boost sensitivity
enrolment <- 0

# Log variables
put(vars_content_to_string(cut, study_start_date, study_end_date, start_yr, age_start, age_end, mem_days, presvcstart, pre_conception_days, enrolment))

# Store file locally (TRUE / FALSE)
data_extraction_save_locally <- TRUE

# Connecting to database -------------

# Setup connection to Impala - variables are present in environment

db_connection <- dbConnect(
  odbc::odbc(),
  driver = Sys.getenv("RDIP_IMPALA_DRIVER"),
  host = Sys.getenv("RDIP_IMPALA_SERVER"),
  port = Sys.getenv("RDIP_IMPALA_PORT"),
  authmech = Sys.getenv("RDIP_IMPALA_AUTH_MECH"),
  ssl = Sys.getenv("RDIP_IMPALA_SSL"),
  trustedcerts = Sys.getenv("RDIP_IMPALA_TRUSTED_CERTS"),
  database = Sys.getenv("RDIP_IMPALA_DATABASE"),
  uid = Sys.getenv("RDIP_IMPALA_USERID"),
  pwd = Sys.getenv("RDIP_IMPALA_PASSWORD")
)

# Show different schemas
#dbGetQuery(db_connection, 'show schemas')

# Create a lazy tbl from the Impala tables

clnprw_marketscan_inpat <- db_connection %>% tbl(in_schema("clnprw_truven_commercial_claims", paste0("t_s_inpat_", cut)))
clnprw_marketscan_outpat <- db_connection %>% tbl(in_schema("clnprw_truven_commercial_claims", paste0("t_o_outpt_", cut)))

# Save pregnancy codes (ICD-CM-9, ICD-CM.CM, and CPT) to database and import lazily
icd9_diagnosis_pregnancy_codes <- write_to_impala_then_read_lazy_table(tablename = "icd9_dx_preg", pregfindr::icd9_diagnosis_pregnancy_codes,
                                                                       schema = "clnprw_rwd", connection = db_connection, overwrite = FALSE)
icd10_diagnosis_pregnancy_codes <- write_to_impala_then_read_lazy_table(tablename = "icd10_dx_preg", pregfindr::icd10_diagnosis_pregnancy_codes,
                                                                        schema = "clnprw_rwd", connection = db_connection, overwrite = FALSE)
icd9_procedure_pregnancy_codes <- write_to_impala_then_read_lazy_table(tablename = "icd9_proc_preg", pregfindr::icd9_procedure_pregnancy_codes,
                                                                       schema = "clnprw_rwd", connection = db_connection, overwrite = FALSE)
icd10_procedure_pregnancy_codes <- write_to_impala_then_read_lazy_table(tablename = "icd10_proc_preg", pregfindr::icd10_procedure_pregnancy_codes,
                                                                        schema = "clnprw_rwd", connection = db_connection, overwrite = FALSE)
cpt_pregnancy_codes <- write_to_impala_then_read_lazy_table(tablename = "cpt_preg", pregfindr::cpt_pregnancy_codes,
                                                            schema = "clnprw_rwd", connection = db_connection, overwrite = FALSE) 

# Identify patients using Inpatient & Outpatient claims -----------------
# 1) We will identify all records for eligible women in the inpatient dataset, then retrieve all the procedures
# they had by matching their procedures with several sets of codes
# 2) We will repeat step 1 for the outpatient records
# 3) We will merge all of these into a long list
# 4) We will standardise age variables 

# Setting time variables
## Start of study period minus a certain number of days to capture pregnancy start dates
pre_service_date <- as.POSIXct(as.character(lubridate::ymd(service_start_date) - lubridate::days(presvcstart)), tz = "UTC")
## End of study period
service_end_date <- as.POSIXct(service_end_date, tz = "UTC")

# 1) Inpatient claims

## Identify all eligible women in inpatient database
#' Note: if slow, pre-selecting vars in .data might speed it up

inpatient_claims <- filter_eligible_women(
  .data = clnprw_marketscan_inpat,
  .age_start = age_start,
  .age_end = age_end,
  .last_yr = last_yr,
  .presvcdate = pre_service_date,
  .svcend = service_end_date,
  .claims_file = "I",
  .variables_to_keep = c(enrolid, svcdate, dobyr, dx1, dx2, dx3, dx4, dxver, pdx, pproc, proc1, claims_file)
)

## Inpatient ICD-9 Dx Codes
icd9_pregnancy_diagnosis_inpatient_claims <- find_matches_in_ICD(
  .data = inpatient_claims,
  inpat = TRUE,
  .dxver = "9",
  .icd_data = icd9_diagnosis_pregnancy_codes,
  .icd_value = "icd9cm",
  .variables_to_keep = c(enrolid, dobyr, dx, description, pdx, claims_file, svcdate, final_trimester_code, outcome, secondary_outcome)
)
## Inpatient ICD-10 Dx Codes
icd10_pregnancy_diagnosis_inpatient_claims <- find_matches_in_ICD(
  .data = inpatient_claims,
  inpat = TRUE,
  .dxver = "0",
  .icd_data = icd10_diagnosis_pregnancy_codes,
  .icd_value = "icd10cm",
  .variables_to_keep = c(enrolid, dobyr, dx, description, pdx, claims_file, svcdate, final_trimester_code, outcome, secondary_outcome)
)

## Inpatient ICD-9 Proc Codes
icd9_pregnancy_procedure_inpatient_claims <- match_on_pregnancy_procedure_code(
  .data = inpatient_claims,
  .proc_data = icd9_procedure_pregnancy_codes,
  .by_var_claims = "pproc",
  .by_var_proc_code = "icd9_proc",
  .variables_to_keep = c(enrolid, dobyr, proc, description, claims_file, svcdate, final_trimester_code, outcome)
)

## Inpatient ICD-10 Proc Codes
icd10_pregnancy_procedure_inpatient_claims <- match_on_pregnancy_procedure_code(
  .data = inpatient_claims,
  .proc_data = icd10_procedure_pregnancy_codes,
  .by_var_claims = "pproc",
  .by_var_proc_code = "icd10_proc",
  .variables_to_keep = c(enrolid, dobyr, proc, description, claims_file, svcdate, final_trimester_code, outcome)
)

## Inpatient CPT Codes
cpt_pregnancy_inpatient_claims <- match_on_pregnancy_cpt_code(
  .data = inpatient_claims,
  .cpt_data = cpt_pregnancy_codes,
  .variables_to_keep = c(enrolid, dobyr, cpt, description, claims_file, svcdate, final_trimester_code, outcome, secondary_outcome)
)

# 2) Outpatient claims

## Identify all eligible women in outpatient dataset
# ID women that fit inclusion criteria of age and time period
outpatient_claims <- filter_eligible_women(
  .data = clnprw_marketscan_outpat,
  .variables_to_keep = c(enrolid, svcdate, dobyr, dx1, dx2, dx3, dx4, dxver, proc1, claims_file),
  .age_start = age_start,
  .age_end = age_end,
  .last_yr = last_yr,
  .presvcdate = pre_service_date,
  .svcend = service_end_date,
  .claims_file = "O"
)

## Outpatient ICD-9 Dx Codes
icd9_pregnancy_diagnosis_outpatient_claims <- find_matches_in_ICD(
  .data = outpatient_claims,
  inpat = FALSE,
  .dxver = "9",
  .icd_data = icd9_diagnosis_pregnancy_codes,
  .icd_value = "icd9cm",
  .variables_to_keep = c(enrolid, dobyr, dx, description, pdx, claims_file, svcdate, final_trimester_code, outcome, secondary_outcome)
)

## Outpatient ICD-10 Dx Codes
icd10_pregnancy_diagnosis_outpatient_claims <- find_matches_in_ICD(
  .data = outpatient_claims,
  inpat = FALSE,
  .dxver = "0",
  .icd_data = icd10_diagnosis_pregnancy_codes,
  .icd_value = "icd10cm",
  .variables_to_keep = c(enrolid, dobyr, dx, description, pdx, claims_file, svcdate, final_trimester_code, outcome, secondary_outcome)
)

## Outpatient CPT Codes
cpt_pregnancy_outpatient_claims <- match_on_pregnancy_cpt_code(
  .data = outpatient_claims,
  .cpt_data = cpt_pregnancy_codes,
  .variables_to_keep = c(enrolid, dobyr, cpt, description, claims_file, svcdate, final_trimester_code, outcome, secondary_outcome)
)
# 3) and 4) Bring datasets into work library to manipulate and clean

## ICD-9/ICD-10: Combine inpatient and outpatient ICD-CM-9 and ICD-CM-10 diagnosis cohorts
sep("Combine inpatient and outpatient ICD-CM-9 diagnosis cohorts")
icd9_pregnancy_diagnoses <- union(
  icd9_pregnancy_diagnosis_inpatient_claims,
  icd9_pregnancy_diagnosis_outpatient_claims
) %>% collect()
put("Step completed successfully.")

sep("Combine inpatient and outpatient ICD-CM-10 diagnosis cohorts")
icd10_pregnancy_diagnoses <- union(
  icd10_pregnancy_diagnosis_inpatient_claims,
  icd10_pregnancy_diagnosis_outpatient_claims
) %>% collect()
put("Step completed successfully.")

sep("Combine inpatient and outpatient ICD-CM-9 and ICD-CM-10 diagnosis cohorts")
icd_pregnancy_diagnosis_union <-
  union_all(
    icd9_pregnancy_diagnoses,
    icd10_pregnancy_diagnoses
  )
put("Step completed successfully.")

## PROC: Combine inpatient procedure datasets
sep("Combine inpatient procedure cohorts")
icd_pregnancy_procedure_union <- union(
  icd9_pregnancy_procedure_inpatient_claims,
  icd10_pregnancy_procedure_inpatient_claims
) %>% collect()
put("Step completed successfully.")

## CPT: Merge inpatient and outpatient CPT code cohorts
sep("Combine inpatient and outpatient CPT code cohorts")
cpt_pregnancy_union <-
  union(
    cpt_pregnancy_inpatient_claims,
    cpt_pregnancy_outpatient_claims
  ) %>% collect()
put("Step completed successfully.")


## Combine ICD9/ICD10, PROC, and CPT into a single dataset and add new age variables
sep("Merge all data into one data set")
patients <-
  bind_rows(
    ## ICD9/ICD10
    icd_pregnancy_diagnosis_union %>%
      rename(code = dx),
    ## PROC
    icd_pregnancy_procedure_union %>%
      rename(code = proc),
    ## CPT
    cpt_pregnancy_union %>%
      rename(code = cpt)
  ) %>%
  calculate_age_variables()

put("Step completed successfully.")

dbDisconnect(db_connection)

# Saving results Locally -------
if (data_extraction_save_locally) {
  sep("Saving results locally")
  arrow::write_parquet(patients, "extraction_result.parquet")
  if (file.exists("extraction_result.parquet")) {
    put("Step completed successfully.")
  } else {
    put("Step failed.")
  }
}

# Closing log -------------
log_close()

# Garbage collection -------
gc()
```

